---
title: "Models and effect sizes"
author: "Camille Crapart"
date: "16 2 2022"
output: 
   bookdown::html_document2:
     code_folding: hide
     toc: true
     toc_float: true
     number_sections: true
     fig_caption: true
bibliography: C:\\Users\\raine\\Documents\\UiO\\Bibtex\\Finstad.bib
link-citations: yes

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F, error = F, fig.align = "center")
options(knitr.kable.NA="", knitr.kable.format ="html")
```

```{r libraries}
library(dplyr)
library(kableExtra)

source("f_plotspatial.R")
fennoscandia <- readRDS("fennoscandia.Rdata")
norge <- fennoscandia %>% filter(nation == "Norway")
```
# Spatial autocorrelation and variance inflation factor

The spatial autocorrelation of the independent variables is evaluated thanks to the Moran's I index. X is the variable of interest, N the number of spatial units, and $w_{ij}$ a diagonal weight matrix.

$$
I = \frac{N}{\Sigma_i \Sigma_j w_{ij}} \frac{\Sigma_i \Sigma_j w_{ij} (X_i-\bar{X})(X_j-\bar{X})}{\Sigma_i (X_i-\bar{X})^2}
$$

Moreover, the collinearity between the independent variables is assessed by using the Variance Inflation Factor (VIF). It is obtained by regressing each independent variables towards the others, and calculated using the resulting $R^2$.  

$$
VIF = \frac{1}{1-R_i^2}
$$

# Models

Three statistical models are used in this article.

## Linear model LM

We first fit a multiple linear regression model, of the form:

$$
y_i = \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + ... + \varepsilon_i
$$

The regression coefficients are easy to interpret as the effect of one predictor on the dependent variable, holding all the other predictors fixed @Casella2013. It is also the model that requires the least computing. It was fitted with the base package in R. 

## Spatial error model SELM

The spatial error model is a linear model in which the spatial autocorrelation is accounted for in the error term. More accurately, it computes the maximum likelihood estimation of the following model @Bivand2021: 
$$y = X \beta + u$$

Where $u$ is the error vector, having the form:

$$u = \rho Mu +\varepsilon$$

\rho is the scalar spatial autoregressive parameter. Mu is a vector of observations on the spatially lagged vector of residuals, defined thanks to the weight matrix M?

The computation was made on R with the "spatialreg" package @Bivand2019. 


## Geographically weigthed regression GWR

Finally, we fitted a geographically weighted regression model, using the GWmodel package in R @Lu2021. A developement of the model using GWmodel is presented in @Gollini2015b. A geographically weighted model computes a linear regression for each point of the dataset, giving more weight to the closest neighbours. The weight of each neighbours is defined by a weight matrix, specific to each model point. Here we chose a decreasing kernel gaussian weighting function. Instead of using an automatically selected bandwith (the bandwith parameter controls the rate of decay of the function), we defined it as the 100 closest neighbours in order to remain consistent with bandwith chosen for the spatial error model. 

# Effect size {#effect-size}

The effect size of each independent variable is expressed as the relative change of TOC depending on a given change of the independent variable. The conversion from model estimate to effect size is detailed below.

* x is the independent variable
* $\hat{\theta}$ is the regression coefficient estimate
* $h(\hat{\theta})$ is the effect size


For the non-transformed independent variables (NDVI,percentage of bogs and arable land, N-deposition), the effect size was calculated taking into account that the dependent variable was logged. 

$\ln ([TOC]_1) = \hat{\theta} x_1$

$\ln ([TOC]_2) = \hat{\theta} x_2$

$\ln ([TOC]_2) - \ln ([TOC]_1) = \hat{\theta}  (x_2 - x_1)$

$\ln \frac{[TOC]_2}{[TOC]_1} = \hat{\theta} \Delta x$

$\frac{[TOC]_2}{[TOC]_1} = exp({\hat{\theta} \Delta x })$


 The delta used for NDVI, bogs and arable land was 0.1 (NDVI is indicated on a scale from 0 to 1 and "bogs", "arable" and "forest" are proportions): 

$h(\hat{\theta}) = exp({\hat{\theta} \times 0.1})$

The evaluation of the standard error is based on the delta method @MacKenzie2018.

$Var(h(\hat{\theta})) = \left( \frac{\delta h(\theta)}{\delta \hat{\theta}} \right) ^2 \times Var(\hat{\theta})$

$Std(h(\hat{\theta})) = \sqrt{\left( \frac{\delta h(\theta)}{\delta \hat{\theta}} \right) ^2 \times Var(\hat{\theta})}$

$Std(h(\hat{\theta})) = \frac{\delta h(\hat{\theta})}{\delta \hat{\theta}} \times Std(\hat{\theta})$

$\frac{\delta h(\hat{\theta})}{\delta \hat{\theta}} = 0.1 \times exp({\hat{\theta} \times 0.1})$

$Std(h(\hat{\theta})) = 0.1 \times 10^{\hat{\theta} \times 0.1} \times Std(\hat{\theta})$



The effect size for the logged variables was calculated as the relative change of TOC concentration for a 10% increase of the independent variable. 

$\log([TOC]_1) = \hat{\theta} \times \log(x)$

$\log([TOC]_2) = \hat{\theta} \times \log(x \times 1.1) = \hat{\theta} \times \log(x) + \hat{\theta} \times \log(1.1)$

$\log([TOC]_2) - \log([TOC]_1) = \hat{\theta} \times \log(1.1)$

$\log([TOC]_2/[TOC]_1) = \hat{\theta} \times \log(1.1) = \log(1.1^{\hat{\theta}})$

$\frac{[TOC]_2}{[TOC]_1} = 1.1^{\hat{\theta}}$

The corresponding standard error is: 

$Var(h(\hat{\theta})) = \left( \frac{\delta h(\theta)}{\delta \hat{\theta}} \right) ^2 \times Var(\hat{\theta})$

$Std(h(\hat{\theta})) = \sqrt{ \left( \frac{\delta h(\hat{\theta})}{\delta \hat{\theta}} \right) ^2 \times Var(\hat{\theta})}$

$Std(h(\hat{\theta})) = \frac{\delta h(\hat{\theta})}{\delta \hat{\theta}} \times Std(\hat{\theta})$

$\frac{\delta h(\hat{\theta})}{\delta \hat{\theta}} = ln(1.1) \ times 1.1^{\hat{\theta}}$

$Std(h(\hat{\theta})) = ln(1.1) \times 1.1^{\hat{\theta}} \times Std(\hat{\theta})$

# Model results

## Summary table LM and SELM Fennoscandia

```{r lm-table-results}
summary.table.lm <- readRDS("summary.table.lm.rds")
summary.table.sem <- readRDS("summary.table.sem.rds")

n <- 6

knitr::kable(summary.table.lm,digits = 3) %>%
  add_header_above(c("Model"=1,"Scaled LM" = 2, "LM" = 2, "Effect size" = 3,"SIM" = 2),italic = T) %>%
  column_spec(column = 8, bold = T) %>% 
  kable_styling(bootstrap_options = "bordered") %>%
  group_rows(group_label = "Model Evaluation", start_row = n+1,end_row=n+3)

knitr::kable(summary.table.sem,digits = 4) %>%
  add_header_above(c("Model"=1,"Scaled SELM" = 2, "SELM" = 2, "Effect size" = 3),italic = T) %>%
  kable_styling(bootstrap_options = "bordered") %>%
  column_spec(column = 8, bold = T) %>% 
  group_rows(group_label = "Model Evaluation", start_row = n+1,end_row=n+2)
```

## Summary tables simple model Norway

```{r summary-table-norway}
summary.table.sem.norge <- readRDS("summary.table.sem.norge.rds")
n <- 3

knitr::kable(summary.table.sem.norge,digits = 4) %>%
  add_header_above(c("Model"=1,"Scaled SELM" = 2,"SELM" = 2, "Effect size" = 3),italic = T) %>%
  kable_styling(bootstrap_options = "bordered") %>%
  column_spec(column = 8, bold = T) %>% 
  group_rows(group_label = "Model Evaluation", start_row = n+1,end_row=n+2)
```

## Summary tables simple model Fennoscandia

```{r summary-table-simple-fen}
summary.table.simple.sem <- readRDS("summary.table.simple.sem.rds")
n <- 3

knitr::kable(summary.table.simple.sem,digits = 4) %>%
  add_header_above(c("Model"=1,"Scaled SELM" = 2,"SELM" = 2, "Effect size" = 3),italic = T) %>%
  kable_styling(bootstrap_options = "bordered") %>%
  column_spec(column = 8, bold = T) %>% 
  group_rows(group_label = "Model Evaluation", start_row = n+1,end_row=n+2)
```

## Residual plots

```{r residual-plots}
# lm models: s.lm.fennoscandia and lm.fennoscandia

s.lm.fennoscandia <- readRDS("s.lm.fennoscandia.rds")
lm.fennoscandia <- readRDS("lm.fennoscandia.rds")

s.sem.fennoscandia <- readRDS("s.sem.fennoscandia.Rdata")
sem.fennoscandia <- readRDS("sem.fennoscandia.Rdata")

sem.norge.simple <- readRDS("sem.norge.simple.Rdata")
sem.fen.simple <- readRDS("sem.fen.simple.Rdata")

sem.forest.fen <- readRDS("sem.forest.fen.Rdata")
sem.forest.norge <- readRDS("sem.forest.norge.Rdata")

f_plotspatial(data = fennoscandia,var = s.lm.fennoscandia$residuals, plottitle = "Residuals for scaled LM")
f_plotspatial(data = fennoscandia,var = s.sem.fennoscandia$residuals, plottitle = "Residuals for scaled SELM")
f_plotspatial(data = fennoscandia,var = lm.fennoscandia$residuals, plottitle = "Residuals for LM")
f_plotspatial(data = fennoscandia,var = sem.fennoscandia$residuals, plottitle = "Residuals for SELM")
f_plotspatial(data = fennoscandia,var = sem.fen.simple$residuals, plottitle = "Residuals for SELM with ndvi, runoff et bogs for Fennoscandia")
f_plotspatial(data = norge,var = sem.norge.simple$residuals, plottitle = "Residuals for SELM with ndvi, runoff et bogs for Norway")
f_plotspatial(data = fennoscandia,var = sem.forest.fen$residuals, plottitle = "Residuals for SELM with forest")
f_plotspatial(data = norge,var = sem.forest.norge$residuals, plottitle = "Residuals for SELM with forest for Norway")

```

## GWR additional plots

```{r explore-bogs}
gwr.effectsize.percent <- readRDS("gwr.ndvi.effectsize.percent.rds")
plot(fennoscandia$bogs*100,gwr.effectsize.percent$bogs, xlab = "Proportion of bogs, %", ylab = "Effect size of a 10% increase in bogs, %")
```

```{r explore-forest}
gwr.effectsize.percent <- readRDS("gwr.forest.effectsize.percent.rds")
plot(fennoscandia$forest*100,gwr.effectsize.percent$forest, xlab = "Proportion of forest, %", ylab = "Effect size of a 10% increase in forest, %")
```


# References
